---
title: "Final Simulation Results"
author: "Corbin Cerny"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Load packages
```{r}
library(dplyr)
library(stringr)
library(readxl)
library(ggplot2)
library(purrr)
library(tidyr)
library(geofacet)
library(tidyverse)
library(purrr)
library(kableExtra)

set.seed(809)

```

# Construct Voter Distributions for Ohio Senate Districts, 2012-2022
To conduct election simulations, data from the Ohio Senate Districts for 2012 thru 2022 is used to construct voter distributions for each district to sample votes from. The variance, density, and median ideological score are used to infer likely distributions of voters by district. This methodology assumes the ideological preferences of a district electorate is reflected by the ideology of all the candidates who stand in elections, not just the winning candidate.

### Load & Prepare the data from DIME (Database on Ideology, Money in Politics, and Elections).

The DIME database includes all candidates and their corresponding ideological scores (`_cfscore_` & `_cfscores.dyn_`).

Subseting into the markers of interest. Namely, we are interested in all state senate candidates ("state:upper") corresponding to the Ohio General Assembly district map in use from 2012 to 2022. Additionally, candidates are ideologically grouped for the purposes of the analysis, with candidates from the GOP and Libertarian parties under the conservative ideology ;abel, and candidates from the Democrat and Green parties under the liberal ideology label.
```{r}
# Load DIME Data
load("C:/Corbin_Documents/Hertie, Class of 2026/Course Work/Theses/2024-25/Simulations/Ohio Distributions/DIME/dime_recipients_1979_2024.rdata")

# Subset to Ohio State Senate Candidates for 2012 thru 2022

cands_oh <- cands |> 
  filter(
    state == "OH",
    cycle >= 2012 & cycle <= 2022,
    seat == "state:upper",
    !is.na(recipient.cfscore.dyn)
  ) |> 
  mutate(
    district = ifelse(str_detect(district, "^OH-\\d+$"), district, paste0("OH-", district)), # Format
    ideological_block = case_when(
      party %in% c(100, 101) ~ "Liberal",       # Democrat, Green
      party %in% c(200, 201) ~ "Conservative",  # Republican, Libertarian
      TRUE ~ NA_character_
    )
  ) |> 
  filter(!is.na(ideological_block))  # Remove any unclassified parties

# Check distribution
table(cands_oh$ideological_block)

# Subset into individual senate districts

district_list <- setNames(
  lapply(1:33, function(i) {
    district_code <- paste0("OH-", i)
    cands_oh[cands_oh$district == district_code, ]
  }),
  paste0("oh", 1:33)
)

```
Statewide, we see that there are a greater number of conservative candidates who ran for state senate seats during the 2012-2022 period than liberal.

The variable _cfscore_ is the estimated ideology of a candidate/recipient while the variable _cfscores.dyn_ is the period-specific estimates of ideology. Under the _.dyn_ scoring, candidate/recipient scores are re-estimated in each election cycle while holding contributor scores constant.

- CFscore is a continuous measure of political ideology derived from campaign finance data
- Higher positive values = More Conservative
- Lower negative values = More Liberal
- 0 represents a centrist or moderate position.
- Candidates, donors, and organizations all receive CFscores, making it possible to compare their ideological positions directly.
- While there is no fixed range for CFscores, they typically fall within the -2 to +2 interval.

### Calculate Variance for Conservative & Liberal Voters Across State
The data we are interested in corresponds to the 2012-2022 General Assembly district map. This ten year period is inclusive of, on average, 3 state senate elections for each district, where odd number districts (ie. OH-1) and even number districts (ie. OH-2) oscillate holding elections every 2 years. This limits the amount of relevant data corresponding to the district map of interest to calculate the variance of each district. Therefore, the variance is calculated for the overall available data, that is, statewide for conservatives and liberals.
```{r}
# Compute variance and standard deviation
ideology_variance <- cands_oh |> 
  group_by(ideological_block) |> 
  summarise(
    variance = var(recipient.cfscore.dyn, na.rm = TRUE),
    sd = sd(recipient.cfscore.dyn, na.rm = TRUE),
    n = n()
  )

print(ideology_variance)

# Create Objects
sd_conservative <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
sd_liberal <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]
```
The results indicate that liberal candidates statewide are more ideologically dispersed than conservative candidates, suggesting that Democratic/Green candidates vary more widely in their ideological positions than Republican/Liberatian candidates, who are more ideologically clustered.

### Calculate Median _cfscores.dyn_ For Conservative & Liberal By District
Calculations use the dynamic CFscore, `_cfscores.dyn_`,  which reflecrs the CFscore of each election cycle in the period of interest. The _.dyn_ scores of all conservative (GOP and Libertarian) and liberal (Democratic and Green) candidates who ran for office between 2012 and 2022 in each district are considered. The median is taken to compute an average ideological score for conservative and liberal voters in each district. Calculating the median within each group helps represent the core ideological stance of each party, even if a few candidates are outliers.
```{r}
# Create empty lists to store medians
median_conservative <- list()
median_liberal <- list()
district_names <- list()  # To store district names

# Loop through the list of districts
for (i in 1:33) {
  district_key <- paste0("oh", i)  # This is the list element name
  district_data <- district_list[[district_key]]  # Access the dataframe
  
  if (!is.null(district_data)) {
    district_names[[district_key]] <- paste0("OH-", i)

    # Compute medians for conservative (>0) and liberal (<0) candidates
    median_conservative[[district_key]] <- median(
      district_data$recipient.cfscore.dyn[district_data$recipient.cfscore.dyn > 0], na.rm = TRUE
    )
    median_liberal[[district_key]] <- median(
      district_data$recipient.cfscore.dyn[district_data$recipient.cfscore.dyn < 0], na.rm = TRUE
    )
  }
}

# Combine into one dataframe
district_medians <- data.frame(
  district = unlist(district_names),
  median_conservative = unlist(median_conservative),
  median_liberal = unlist(median_liberal)
)

```

*In OH-12, DIME is missing data on the lone democrat - Ken Poling who ran in 2020 - to run for state senate there during the whole period of interest (2012 thru 2022). No liberal candidate data exists for OH-12. Therefore, no median score can be calculated for OH-12 liberals.*

### Calculate Density For Conservative & Liberal By District
The share of votes captured in each election from 2012-2022 is used to determine the approximate share of each district electorate is conservative and liberal. The average conservative and liberal vote share is calculated by district using data from each election occurring during the period of interest. 
```{r}
# Load Data
elections <- read_excel("C:/Corbin_Documents/Hertie, Class of 2026/Course Work/Theses/2024-25/Simulations/Ohio Distributions/Ohio Senate Districts - 2012-2022 Election Results.xlsx") |> 
  mutate(district = gsub("District ", "OH-", district))

# Compute average vote share for each district across all elections from 2012 to 2022
avg_vote_share <- elections |> 
  mutate(
    total_votes = democrat.votes + green.votes + republican.votes + libertarian.votes,
    liberal_percent = (democrat.votes + green.votes) / total_votes * 100,
    conservative_percent = (republican.votes + libertarian.votes) / total_votes * 100
  ) |> 
  group_by(district) |> 
  summarise(
    avg_liberal_percent = median(liberal_percent, na.rm = TRUE),
    avg_conservative_percent = median(conservative_percent, na.rm = TRUE)
  )

# Merge with the district_medians dataframe
district_medians <- district_medians |> 
  left_join(avg_vote_share, by = "district")
```


# Simulations - Plurality, TRM, IRV, Condorcet-Borda, Copeland Round-Robin

##Setup
```{r}
# ----- SETTINGS -----
total_voters <- 10000
lambda <- 0.5
n_primary_candidates <- 4
n_simulations <- 100

# ----- SOFTMAX FUNCTION -----
softmax_vote <- function(voters, candidates, lambda) {
  sapply(voters, function(v) {
    u0 <- -abs(v - candidates) / lambda    # divide by λ
    u  <- u0 - max(u0)                     # shift
    p  <- exp(u) / sum(exp(u))
    sample.int(length(p), 1, prob = p)
  })
}

```

### Closed Plurality Election
```{r}
# ----- CLOSED PLURALITY FUNCTION -----
simulate_closed_plurality <- function(
  row,
  total_voters,
  n_primary_candidates,
  iterations,
  lambda
) {
  cat("Running CLOSED primary for", row$district, "\n")

  # skip if medians missing
  if (is.na(row$median_conservative) || is.na(row$median_liberal)) {
    return(NULL)
  }

  # unpack district parameters
  m_con  <- row$median_conservative
  m_lib  <- row$median_liberal
  p_con  <- row$avg_conservative_percent
  p_lib  <- row$avg_liberal_percent

  # bloc dispersions from your global table
  sd_con <- ideology_variance$sd[
    ideology_variance$ideological_block == "Conservative"
  ]
  sd_lib <- ideology_variance$sd[
    ideology_variance$ideological_block == "Liberal"
  ]

  # compute bloc sizes
  con_prop <- p_con / (p_con + p_lib)
  n_con    <- round(total_voters * con_prop)
  n_lib    <- total_voters - n_con

  # Monte Carlo
  replicate(iterations, {
    # 1) draw voter ideals
    con_voters <- rnorm(n_con, mean = m_con, sd = sd_con)
    lib_voters <- rnorm(n_lib, mean = m_lib, sd = sd_lib)

    # 2) conservative primary
    con_cands  <- rnorm(n_primary_candidates, mean = m_con, sd = sd_con)
    con_ballot <- softmax_vote(con_voters, con_cands, lambda)
    con_winner <- con_cands[ which.max(
      table(factor(con_ballot, levels = seq_along(con_cands)))
    ) ]

    # 3) liberal primary
    lib_cands  <- rnorm(n_primary_candidates, mean = m_lib, sd = sd_lib)
    lib_ballot <- softmax_vote(lib_voters, lib_cands, lambda)
    lib_winner <- lib_cands[ which.max(
      table(factor(lib_ballot, levels = seq_along(lib_cands)))
    ) ]

    # 4) general election
    all_voters <- c(con_voters, lib_voters)
    finalists  <- c(con_winner, lib_winner)
    gen_ballot <- softmax_vote(all_voters, finalists, lambda)
    winner_idx <- which.max(table(factor(gen_ballot, 1:2)))

    finalists[winner_idx]
  })
}

```

### Two-Round Majority (TRM) Election
```{r}

# ----- TRM FUNCTION -----
simulate_trm <- function(row, iterations = 100) {
  print(glue::glue("Running {row$district}"))
  if (is.na(row$median_conservative) || is.na(row$median_liberal)) return(NULL)

  # Load parameters
  median_conservative <- row$median_conservative
  median_liberal <- row$median_liberal
  avg_conservative_percent <- row$avg_conservative_percent
  avg_liberal_percent <- row$avg_liberal_percent

  sd_conservative <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
  sd_liberal <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]

  con_prop <- avg_conservative_percent / (avg_conservative_percent + avg_liberal_percent)
  lib_prop <- 1 - con_prop

  n_conservative <- round(total_voters * con_prop)
  n_liberal <- total_voters - n_conservative

  replicate(iterations, {
    con_voters <- rnorm(n_conservative, mean = median_conservative, sd = sd_conservative)
    lib_voters <- rnorm(n_liberal, mean = median_liberal, sd = sd_liberal)
    all_voters <- c(con_voters, lib_voters)

    electorate_mean <- weighted.mean(c(median_conservative, median_liberal), w = c(avg_conservative_percent, avg_liberal_percent))
    electorate_sd <- weighted.mean(c(sd_conservative, sd_liberal), w = c(avg_conservative_percent, avg_liberal_percent))
    primary_candidates <- rnorm(n_primary_candidates, mean = electorate_mean, sd = electorate_sd)

    primary_choices <- softmax_vote(all_voters, primary_candidates, lambda)
    primary_votes <- table(factor(primary_choices, levels = 1:n_primary_candidates))
    top_two_indices <- order(primary_votes, decreasing = TRUE)[1:2]
    general_candidates <- primary_candidates[top_two_indices]

    general_choices <- softmax_vote(all_voters, general_candidates, lambda)
    general_votes <- table(factor(general_choices, levels = 1:2))
    winner_index <- which.max(general_votes)

    general_candidates[winner_index]
  })
}

```

### IRV Election
```{r}

# ----- IRV FUNCTION -----
simulate_irv <- function(row) {
  cat("Running IRV for", row$district, "\n")
  
  # unpack district parameters
  m_con  <- row$median_conservative
  m_lib  <- row$median_liberal
  p_con  <- row$avg_conservative_percent
  p_lib  <- row$avg_liberal_percent
  district <- row$district
  
  if (is.na(m_con) || is.na(m_lib)) return(NULL)
  
  # bloc dispersions
  sd_con <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
  sd_lib <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]
  
  # run N simulations
  replicate(n_simulations, {
    # 1) draw voter ideals
    con_prop     <- p_con / (p_con + p_lib)
    n_con        <- round(total_voters * con_prop)
    n_lib        <- total_voters - n_con
    con_voters   <- rnorm(n_con, mean = m_con, sd = sd_con)
    lib_voters   <- rnorm(n_lib, mean = m_lib, sd = sd_lib)
    all_voters   <- c(con_voters, lib_voters)
    
    # 2) draw candidate positions
    electorate_mean <- weighted.mean(c(m_con, m_lib), w = c(p_con, p_lib))
    electorate_sd   <- weighted.mean(c(sd_con, sd_lib), w = c(p_con, p_lib))
    candidates      <- rnorm(n_primary_candidates,
                             mean = electorate_mean,
                             sd   = electorate_sd)
    
    # 3) build *softmax*‐based full rankings
    #    each voter i: iteratively sample next‐best from remaining
    ballots <- t(sapply(all_voters, function(v) {
      remaining_cands  <- candidates
      remaining_idxs   <- seq_along(candidates)
      ranking          <- integer(length(candidates))
      
      for (pos in seq_along(ranking)) {
        # pick next favorite among remaining
        pick_rel_index <- softmax_vote(v, remaining_cands, lambda)
        # map back to original index
        pick_orig      <- remaining_idxs[pick_rel_index]
        ranking[pos]   <- pick_orig
        
        # drop that candidate
        remaining_cands <- remaining_cands[-pick_rel_index]
        remaining_idxs  <- remaining_idxs[-pick_rel_index]
      }
      
      ranking
    }))
    
    # 4) IRV elimination
    remaining <- seq_len(n_primary_candidates)
    repeat {
      # first‐choice counts among remaining
      first_choices <- apply(ballots, 1, function(rank) {
        # pick highest‐ranked still in 'remaining'
        rank[ rank %in% remaining ][1]
      })
      counts <- table(factor(first_choices, levels = remaining))
      
      # winner if >50% or only one left
      if (any(counts > length(all_voters) / 2) || length(remaining) == 1) {
        win_idx <- as.numeric(names(which.max(counts)))
        return(data.frame(district = district,
                          winner_ideology = candidates[win_idx]))
      }
      
      # otherwise, eliminate the lowest
      min_count <- min(counts)
      elim      <- as.numeric(names(counts)[counts == min_count])[1]
      remaining <- setdiff(remaining, elim)
    }
    
  }, simplify = FALSE) %>% bind_rows()
}


```

### Condorcet - Borda
```{r}
# ----- CONDORCET - BORDA SIMULATION FUNCTION -----
simulate_condorcet <- function(row) {
  cat("Running Condorcet-Borda for", row$district, "\n")
  
  # unpack district parameters
  m_con  <- row$median_conservative
  m_lib  <- row$median_liberal
  p_con  <- row$avg_conservative_percent
  p_lib  <- row$avg_liberal_percent
  district <- row$district
  
  if (is.na(m_con) || is.na(m_lib)) return(NULL)
  
  # bloc dispersions
  sd_con <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
  sd_lib <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]
  
  replicate(n_simulations, {
    # 1) draw voter ideals
    con_prop   <- p_con / (p_con + p_lib)
    n_con      <- round(total_voters * con_prop)
    n_lib      <- total_voters - n_con
    con_voters <- rnorm(n_con, mean = m_con, sd = sd_con)
    lib_voters <- rnorm(n_lib, mean = m_lib, sd = sd_lib)
    all_voters <- c(con_voters, lib_voters)
    
    # 2) draw candidate positions
    electorate_mean <- weighted.mean(c(m_con, m_lib), w = c(p_con, p_lib))
    electorate_sd   <- weighted.mean(c(sd_con, sd_lib), w = c(p_con, p_lib))
    candidates      <- rnorm(n_primary_candidates,
                             mean = electorate_mean,
                             sd   = electorate_sd)
    
    # 3) build *softmax*‐based full rankings (for Borda fallback)
    ballots <- t(sapply(all_voters, function(v) {
      remaining    <- candidates
      remaining_ix <- seq_along(candidates)
      ranking      <- integer(length(candidates))
      
      for (pos in seq_along(ranking)) {
        # sample next favorite among remaining
        pick_rel <- softmax_vote(v, remaining, lambda)
        pick_orig <- remaining_ix[pick_rel]
        ranking[pos] <- pick_orig
        
        # drop chosen
        remaining    <- remaining[-pick_rel]
        remaining_ix <- remaining_ix[-pick_rel]
      }
      
      ranking
    }))
    
    # 4) pairwise softmax‐based Condorcet matrix
    win_matrix <- matrix(0, n_primary_candidates, n_primary_candidates)
    for (i in 1:n_primary_candidates) {
      for (j in seq_len(n_primary_candidates)) {
        if (i == j) next
        
        # for each voter, softmax vote between just candidates i & j
        picks <- softmax_vote(all_voters,
                              c(candidates[i], candidates[j]),
                              lambda)
        # count how many prefer i (pick==1)
        wins_i <- sum(picks == 1)
        if (wins_i > length(all_voters) / 2) {
          win_matrix[i, j] <- 1
        }
      }
    }
    
    # 5) pick Condorcet winner if one exists
    cond_winner <- which(rowSums(win_matrix) == (n_primary_candidates - 1))
    if (length(cond_winner) == 1) {
      return(data.frame(district = district,
                        winner_ideology = candidates[cond_winner]))
    }
    
    # 6) otherwise Borda‐count fallback
    #    assign Borda points based on the softmax rankings
    borda_scores <- integer(n_primary_candidates)
    for (v in seq_len(nrow(ballots))) {
      # higher rank (pos=1) gets n-1 points, pos=2 gets n-2, etc.
      ranks <- ballots[v, ]
      borda_scores[ranks] <- borda_scores[ranks] +
                             (n_primary_candidates - match(ranks, ranks))
    }
    fallback    <- which.max(borda_scores)
    data.frame(district = district,
               winner_ideology = candidates[fallback])
    
  }, simplify = FALSE) %>% bind_rows()
}

```

### Copeland Round-Robin
```{r}
# ----- COPELAND SIMULATION FUNCTION -----
simulate_copeland <- function(row) {
  cat("Running Copeland for", row$district, "\n")
  
  # unpack district parameters
  m_con  <- row$median_conservative
  m_lib  <- row$median_liberal
  p_con  <- row$avg_conservative_percent
  p_lib  <- row$avg_liberal_percent
  district <- row$district
  
  if (is.na(m_con) || is.na(m_lib)) return(NULL)
  
  # bloc dispersions
  sd_con <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
  sd_lib <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]
  
  replicate(n_simulations, {
    # 1) draw voter ideals
    con_prop   <- p_con / (p_con + p_lib)
    n_con      <- round(total_voters * con_prop)
    n_lib      <- total_voters - n_con
    con_voters <- rnorm(n_con, mean = m_con, sd = sd_con)
    lib_voters <- rnorm(n_lib, mean = m_lib, sd = sd_lib)
    all_voters <- c(con_voters, lib_voters)
    
    # 2) draw candidate positions
    electorate_mean <- weighted.mean(c(m_con, m_lib), w = c(p_con, p_lib))
    electorate_sd   <- weighted.mean(c(sd_con, sd_lib), w = c(p_con, p_lib))
    candidates      <- rnorm(n_primary_candidates,
                             mean = electorate_mean,
                             sd   = electorate_sd)
    
    # 3) build pairwise softmax‐based win matrix
    win_matrix <- matrix(0, n_primary_candidates, n_primary_candidates)
    for (i in seq_len(n_primary_candidates)) {
      for (j in seq_len(n_primary_candidates)) {
        if (i == j) next
        
        # head‐to‐head noisy votes via softmax
        head2head_picks <- softmax_vote(
          all_voters,
          c(candidates[i], candidates[j]),
          lambda
        )
        
        # count how many voters prefer i (code 1) vs j (code 2)
        wins_i <- sum(head2head_picks == 1)
        wins_j <- length(all_voters) - wins_i
        
        if (wins_i > wins_j)      win_matrix[i, j] <-  1
        else if (wins_i < wins_j) win_matrix[i, j] <- -1
        # ties leave 0
      }
    }
    
    # 4) Copeland score & winner
    copeland_scores <- rowSums(win_matrix)
    winner_idx      <- which.max(copeland_scores)
    
    data.frame(
      district        = district,
      winner_ideology = candidates[winner_idx]
    )
    
  }, simplify = FALSE) %>% bind_rows()
}


```

## Conduct Simulated Elections
```{r}
# ----- CLOSED PLURALITY -----
winners_closed_plurality_df <- map_dfr(
  seq_len(nrow(district_medians)),
  function(i) {
    row <- district_medians[i, ]
    wins <- simulate_closed_plurality(
      row,
      total_voters         = total_voters,
      n_primary_candidates = n_primary_candidates,
      iterations           = n_simulations,
      lambda               = lambda
    )
    if (is.null(wins)) return(NULL)
    data.frame(
      district        = row$district,
      winner_ideology = wins
    )
  }
)

# ----- TRM -----
winners_trm_df <- purrr::map_dfr(1:nrow(district_medians), function(i) {
  row <- district_medians[i, ]
  district <- row$district

  winner_ideologies <- simulate_trm(row, iterations = n_simulations)
  if (is.null(winner_ideologies)) return(NULL)

  data.frame(
    district = rep(district, length(winner_ideologies)),
    winner_ideology = winner_ideologies
  )
})


# ----- IRV -----
winners_irv_df <- district_medians %>%
  split(.$district) %>%
  map_dfr(simulate_irv)


# ----- CONDORCET BORDA -----
winners_borda_df <- district_medians %>%
  split(.$district) %>%
  map_dfr(simulate_condorcet)

# ----- COPELAND ROUND ROBIN -----
winners_copeland_df <- district_medians %>%
  split(.$district) %>%
  map_dfr(simulate_copeland)


# Classify Winners As Liberal Or Conservative
winners_closed_plurality_df <- winners_closed_plurality_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"))

winners_trm_df <- winners_trm_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"))

winners_irv_df <- winners_irv_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"))

winners_borda_df <- winners_borda_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"))

winners_copeland_df <- winners_copeland_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"))

```

## Visualize Results
```{r}
# Add class and method labels
add_labels <- function(df, method_name) {
  df %>%
    mutate(
      class = ifelse(winner_ideology > 0, "Conservative", "Liberal"),
      method = method_name
    )
}

winners_closed_plurality_df <- add_labels(winners_closed_plurality_df, "Closed Plurality")
winners_trm_df <- add_labels(winners_trm_df, "TRM")
winners_irv_df       <- add_labels(winners_irv_df, "IRV")
winners_copeland_df  <- add_labels(winners_copeland_df, "Copeland")
winners_borda_df <- add_labels(winners_borda_df, "Condorcet-Borda")

combined_results <- bind_rows(
  winners_closed_plurality_df,
  winners_trm_df,
  winners_irv_df,
  winners_copeland_df,
  winners_borda_df
)

# Calculate medians
median_labels <- combined_results |> 
  group_by(method, class) |> 
  summarise(median = median(winner_ideology, na.rm = TRUE), .groups = "drop") |> 
  pivot_wider(names_from = class, values_from = median) |> 
  mutate(ideological_distance = abs(Conservative - Liberal),
         y_pos = 1.2)  # Choose y-position above the boxes

method_order <- c("Closed Plurality", "TRM", "IRV", "Condorcet-Borda", "Copeland")

# Convert `method` to a factor with specified levels
combined_results$method <- factor(combined_results$method, levels = method_order)
median_labels$method <- factor(median_labels$method, levels = method_order)

ggplot(combined_results, aes(x = winner_ideology, y = class, fill = class)) +
  geom_boxplot(outlier.size = 1, width = 0.5) +
  facet_wrap(~ method, ncol = 1) +
  scale_fill_manual(values = c("Conservative" = "firebrick", "Liberal" = "blue")) +
  labs(
    title = "Ideological Distribution of Election Winners by Voting System",
    x = "Winner Ideology Score",
    y = "Ideological Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

# Median Summary Table
median_labels %>%
  mutate(method = factor(method, levels = method_order)) %>%
  arrange(method) %>%
  select(method, Conservative, Liberal, ideological_distance) %>%
  rename(
    `Voting System` = method,
    `Conservative Median` = Conservative,
    `Liberal Median` = Liberal,
    `Ideological Distance (Δ)` = ideological_distance
  ) %>%
  mutate(across(where(is.numeric), ~ round(., 2))) %>%
  kable("html", caption = "Median Ideologies and Distance by Voting System") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(2:4, bold = TRUE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#555555")


```

*Note on the use of Softmax - The use of softmax in IRV, Borda, and Copeland is a design choice to model individual decision noise, reflecting probabilistic choice. These systems are used to select a single winner per district, where softmax helps simulate realistic uncertainty in voter preferences, without breaking the logic of the system. In multi-winner systems like STV, deterministic and valid full rankings of candidates are needed. Here, softmax introduces ambiguity that breaks the STV logic. Most academic and practitioner simulations of STV in the literature assume voters rank deterministically based on proximity (e.g., ideological distance), which guarantees valid ballots for the quota and elimination logic to work as intended.*

# Simulation - Single Transferable Vote With Multiple Winners

STV is used to determine a slate of winners for multi-winner elections (multi-member districts) rather than single winner (single-member districts), as is the case for plurality, IRV, Condorcet-Borda, and Copeland Round-Robin. Senate districts in Ohio comprise 3 House districts, each electing a single-member. In this simulation scenario, it is theorized that the single-member Senate districts are expanded to a 3-member district (a reflection of the current Senate-House district ratio) and STV is used to select multiple winners to fill the slate of representatives for each district.

```{r}
# ----- STV SETTINGS -----
n_stv_primary_candidates <- 6
n_stv_winners <- 3


# ----- STV FUNCTION -----
simulate_stv <- function(row, lambda, iterations = n_simulations) {
  cat("Running STV for", row$district, "\n")
  district <- row$district
  
  # skip if medians missing
  if (is.na(row$median_conservative) || is.na(row$median_liberal)) return(NULL)

  # unpack parameters
  m_con <- row$median_conservative
  m_lib <- row$median_liberal
  p_con <- row$avg_conservative_percent
  p_lib <- row$avg_liberal_percent

  # bloc dispersions
  sd_con <- ideology_variance$sd[ideology_variance$ideological_block == "Conservative"]
  sd_lib <- ideology_variance$sd[ideology_variance$ideological_block == "Liberal"]

  replicate(iterations, {
    # 1) draw voter ideals
    con_prop   <- p_con / (p_con + p_lib)
    n_con      <- round(total_voters * con_prop)
    n_lib      <- total_voters - n_con
    con_voters <- rnorm(n_con, mean = m_con, sd = sd_con)
    lib_voters <- rnorm(n_lib, mean = m_lib, sd = sd_lib)
    all_voters <- c(con_voters, lib_voters)

    # 2) draw candidate positions
    electorate_mean <- weighted.mean(c(m_con, m_lib), w = c(p_con, p_lib))
    electorate_sd   <- weighted.mean(c(sd_con, sd_lib), w = c(p_con, p_lib))
    candidates      <- rnorm(n_stv_primary_candidates, mean = electorate_mean, sd = electorate_sd)

    # 3) build *softmax*-based full rankings
    ballots <- t(sapply(all_voters, function(v) {
      remaining_cands <- candidates
      remaining_ix    <- seq_along(candidates)
      ranking         <- integer(length(candidates))
      for (pos in seq_along(ranking)) {
        pick_rel  <- softmax_vote(v, remaining_cands, lambda)
        pick_orig <- remaining_ix[pick_rel]
        ranking[pos] <- pick_orig
        remaining_cands <- remaining_cands[-pick_rel]
        remaining_ix    <- remaining_ix[-pick_rel]
      }
      ranking
    }))

    # 4) STV quota and counting with surplus transfer
    n_voters  <- length(all_voters)
    weights   <- rep(1, n_voters)
    quota     <- floor(sum(weights) / (n_stv_winners + 1)) + 1
    remaining <- seq_len(n_stv_primary_candidates)
    winners   <- integer(0)

    # loop until enough winners or no more candidates
    while (length(winners) < n_stv_winners && length(remaining) > 0) {
      # 4a) first-preference counts (weighted)
      first_choices <- apply(ballots, 1, function(b) b[b %in% remaining][1])
      counts <- tapply(weights, first_choices, sum)
      counts <- counts[as.character(remaining)]
      names(counts) <- remaining

      # 4b) elect any above quota
      electees <- as.numeric(names(counts)[counts >= quota])
      if (length(electees) > 0) {
        # transfer surplus for each elected candidate
        for (cand in setdiff(electees, winners)) {
          if (length(winners) < n_stv_winners) {
            votes     <- counts[as.character(cand)]
            surplus   <- votes - quota
            transfer_fraction <- surplus / votes
            # identify ballots to transfer
            idx <- which(first_choices == cand)
            weights[idx] <- weights[idx] * transfer_fraction
            winners <- c(winners, cand)
            remaining <- setdiff(remaining, cand)
          }
        }
        next  # re-count after transfers
      }

      # 4c) eliminate lowest if no one reached quota
      lowest <- as.numeric(names(counts)[counts == min(counts)])[1]
      remaining <- setdiff(remaining, lowest)
    }

    # ensure exactly n_stv_winners
    winners <- winners[1:min(length(winners), n_stv_winners)]
    winner_ideologies <- candidates[winners]

    # return one row per winner
    data.frame(
      district        = rep(district, length(winner_ideologies)),
      winner_ideology = winner_ideologies
    )

  }, simplify = FALSE) |> bind_rows()
}



# ----- RUN STV SIMULATIONS -----
winners_stv_df <- district_medians %>%
  split(.$district) %>%
  map_dfr(~ simulate_stv(.x, lambda, n_simulations))

```

## Visualize Results
```{r}
# Classify Winners As Liberal Or Conservative
winners_stv_df <- winners_stv_df |> 
  mutate(class = if_else(winner_ideology > 0, "Conservative", "Liberal"),
         method = "STV")

# Calculate medians by class
medians_stv <- winners_stv_df %>%
  group_by(class) %>%
  summarise(median_ideology = median(winner_ideology), .groups = "drop")

# Calculate distance between medians
if (nrow(medians_stv) == 2) {
  delta <- abs(diff(medians_stv$median_ideology))
  print(glue::glue("Δ (Distance between medians) = {round(delta, 2)}"))
} else {
  warning("Not enough classes to compute distance.")
}

# Test Plot
ggplot(winners_stv_df, aes(x = winner_ideology, y = class, fill = class)) +
  geom_boxplot(outlier.size = 1, width = 0.5) +
  scale_fill_manual(values = c("Conservative" = "firebrick", "Liberal" = "blue")) +
  labs(
    title = "Ideological Distribution of Election Winners for STV",
    x = "Winner Ideology Score",
    y = "Ideological Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")
```
## Final Table of Median Distances
```{r}

method_order_final <- c("Closed Plurality", "STV", "IRV", "TRM", "Copeland", "Condorcet-Borda")

# Create STV row as a tibble with correct names and order
stv_row <- tibble(
  `Voting System` = "STV",
  `Conservative Median` = round(medians_stv$median_ideology[medians_stv$class == "Conservative"], 2),
  `Liberal Median` = round(medians_stv$median_ideology[medians_stv$class == "Liberal"], 2),
  `Ideological Distance (Δ)` = round(delta, 2)
)

# Combine with existing summary
table_data <- median_labels %>%
  filter(method != "STV") %>%    # <-- remove STV if it already exists
  select(method, Conservative, Liberal, ideological_distance) %>%
  rename(
    `Voting System` = method,
    `Conservative Median` = Conservative,
    `Liberal Median` = Liberal,
    `Ideological Distance (Δ)` = ideological_distance
  ) %>%
  mutate(across(where(is.numeric), round, 2)) %>%
  bind_rows(stv_row) %>%
  mutate(`Voting System` = factor(`Voting System`, levels = method_order_final)) %>%
  arrange(`Voting System`)

# Generate table
table_data %>%
  kable("html", caption = "Median Ideologies and Distance by Voting System") %>%
  kable_styling(full_width = FALSE, bootstrap_options = c("striped", "hover", "condensed", "responsive")) %>%
  column_spec(2:4, bold = TRUE) %>%
  row_spec(0, bold = TRUE, color = "white", background = "#555555")


```

## Final Plot of Election Outcomes
```{r}
# Combine with existing systems
combined_results <- bind_rows(
  winners_closed_plurality_df,
  winners_trm_df,
  winners_irv_df,
  winners_copeland_df,
  winners_borda_df,
  winners_stv_df  # <- Add STV
)

# Calculate medians again with STV included
median_labels <- combined_results |> 
  group_by(method, class) |> 
  summarise(median = median(winner_ideology, na.rm = TRUE), .groups = "drop") |> 
  pivot_wider(names_from = class, values_from = median) |> 
  mutate(ideological_distance = abs(Conservative - Liberal),
         y_pos = 1.2)

# Reorder methods for plot and labels
combined_results$method <- factor(combined_results$method, levels = method_order_final)
median_labels$method <- factor(median_labels$method, levels = method_order_final)

# Final plot with STV included
ggplot(combined_results, aes(x = winner_ideology, y = class, fill = class)) +
  geom_boxplot(outlier.size = 1, width = 0.5) +
  facet_wrap(~ method, ncol = 1) +
  scale_fill_manual(values = c("Conservative" = "firebrick", "Liberal" = "blue")) +
  labs(
    title = "Ideological Distribution of Election Winners by Voting System",
    x = "Winner Ideology Score",
    y = "Ideological Class"
  ) +
  theme_minimal(base_size = 13) +
  theme(legend.position = "bottom")

```


